---
title: "358_RDS_Project"
author: "Jodeci Wheaden"
date: "2023-01-18"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries
```{r libraries}
library(ggplot2)
library(dplyr)
```

## Set Working Directory

```{r setWD}
# set working directory
setwd('/Volumes/GoogleDrive/My Drive/Reproducible Data Science/BST270-Winter2023/individual_project/project')
```

## Global Variables

```{r global_variables}
# Initializing city vector
city = c()
# Initializing No. of Years Vector
no.ofYears = c()
# Initializing Total Vector
Total = c()
# Initializing Yearly Avg vector
YearlyAvg = c()
```

## Atlanta, GA
```{r import_Atlanta_Data}
atlanta_data = read.csv(file = './atlanta_edited.csv')
```

```{r police_settlement_amts_Atlanta}
# Found on Read.me in the List of cities FOIA'd and what happened with each table
number_years_GA = 2020-2015
total_settlement_GA = sum(atlanta_data$amount_awarded)
yearly_average_GA = total_settlement_GA/number_years_GA

# Append elements to vector
city = append(city, "Atlanta, GA")
no.ofYears = append(no.ofYears, number_years_GA)
Total = append(Total, total_settlement_GA)
YearlyAvg = append(YearlyAvg, yearly_average_GA)
```

## Baltimore, MD
```{r import_Baltimore_Data}
baltimore_data = read.csv(file = './baltimore_edited.csv')
```

```{r police_settlement_amts_Baltimore}
number_years_MD = 2020-2015
total_settlement_MD = sum(baltimore_data$amount_awarded)
yearly_average_MD = total_settlement_MD/number_years_MD

# Append elements to vector
city = append(city, "Baltimore, MD")
no.ofYears = append(no.ofYears, number_years_MD)
Total = append(Total, total_settlement_MD)
YearlyAvg = append(YearlyAvg, yearly_average_MD)
```

## Baton Rouge, LA
```{r import_Baton_Rouge_Data}
baton_rouge_data = read.csv(file = './baton_rouge_edited.csv')
```

```{r police_settlement_amts_BatonRouge}
number_years_LA = 2019-2010 + 1
total_settlement_LA = sum(baton_rouge_data$amount_awarded)
yearly_average_LA = total_settlement_LA/number_years_LA

# Append elements to vector
city = append(city, "Baton Rouge, LA")
no.ofYears = append(no.ofYears, number_years_LA)
Total = append(Total, total_settlement_LA)
YearlyAvg = append(YearlyAvg, yearly_average_LA)
```

## Boston, MA
```{r import_Boston_Data}
boston_data = read.csv(file = './boston_edited.csv')
```

```{r police_settlement_amts_Boston}
number_years_MA = 2019-2010 + 1
total_settlement_MA = sum(boston_data$amount_awarded)
yearly_average_MA = total_settlement_MA/number_years_MA

# Append elements to vector
city = append(city, "Boston, MA")
no.ofYears = append(no.ofYears, number_years_MA)
Total = append(Total, total_settlement_MA)
YearlyAvg = append(YearlyAvg, yearly_average_MA)
```

## Cambridge, MA
```{r import_Cambridge_Data}
cambridge_data = read.csv(file = './cambridge_edited.csv')
```

```{r police_settlement_amts_Cambridge}
number_years_MA2 = 2019-2010 + 1
total_settlement_MA2 = sum(cambridge_data$amount_awarded)
yearly_average_MA2 = total_settlement_MA2/number_years_MA2

# Append elements to vector
city = append(city, "Cambridge, MA")
no.ofYears = append(no.ofYears, number_years_MA2)
Total = append(Total, total_settlement_MA2)
YearlyAvg = append(YearlyAvg, yearly_average_MA2)
```

## Charleston, SC
```{r import_Charleston_Data}
charleston_data = read.csv(file = './charleston_edited.csv')
```

```{r police_settlement_amts_Charleston}
number_years_SC = 2019-2010 + 1
total_settlement_SC = sum(charleston_data$amount_awarded)
yearly_average_SC = total_settlement_SC/number_years_SC

# Append elements to vector
city = append(city, "Charleston, SC")
no.ofYears = append(no.ofYears, number_years_SC)
Total = append(Total, total_settlement_SC)
YearlyAvg = append(YearlyAvg, yearly_average_SC)
```

## Chicago, IL
```{r import_Chicago_Data}
chicago_data = read.csv(file = './chicago_edited.csv')
```

```{r police_settlement_amts_Chicago}
number_years_IL = 2019-2010 + 1
total_settlement_IL = sum(chicago_data$amount_awarded)
yearly_average_IL = total_settlement_IL/number_years_IL

# Append elements to vector
city = append(city, "Chicago, IL")
no.ofYears = append(no.ofYears, number_years_IL)
Total = append(Total, total_settlement_IL)
YearlyAvg = append(YearlyAvg, yearly_average_IL)
```

## Cincinnati, OH
```{r import_Cincinnati_Data}
cincinnati_data = read.csv(file = './cincinnati_edited.csv')
```

```{r police_settlement_amts_Cincinnati}
number_years_OH = 2020-2010 + 1
total_settlement_OH = sum(cincinnati_data$amount_awarded)
yearly_average_OH = total_settlement_OH/number_years_OH

# Append elements to vector
city = append(city, "Cincinnati, OH")
no.ofYears = append(no.ofYears, number_years_OH)
Total = append(Total, total_settlement_OH)
YearlyAvg = append(YearlyAvg, yearly_average_OH)
```

## Cleveland, OH
```{r import_Cleveland_Data}
cleveland_data = read.csv(file = './cleveland_edited.csv')
```

```{r police_settlement_amts_Cleveland}
number_years_OH2 = 2020-2010 + 1
total_settlement_OH2 = sum(cleveland_data$amount_awarded)
yearly_average_OH2 = total_settlement_OH2/number_years_OH2

# Append elements to vector
city = append(city, "Cleveland, OH")
no.ofYears = append(no.ofYears, number_years_OH2)
Total = append(Total, total_settlement_OH2)
YearlyAvg = append(YearlyAvg, yearly_average_OH2)
```

## Columbia, SC
```{r import_Columbia_Data}
columbia_data = read.csv(file = './columbia_edited.csv')
```

```{r police_settlement_amts_Columbia}
number_years_SC2 = 2019-2010 + 1
total_settlement_SC2 = sum(columbia_data$amount_awarded)
yearly_average_SC2 = total_settlement_SC2/number_years_SC2

# Append elements to vector
city = append(city, "Columbia, SC")
no.ofYears = append(no.ofYears, number_years_SC2)
Total = append(Total, total_settlement_SC2)
YearlyAvg = append(YearlyAvg, yearly_average_SC2)
```

## Detroit, MI
```{r import_Detroit_Data}
detroit_data = read.csv(file = './detroit_edited.csv')
```

```{r police_settlement_amts_Detroit}
number_years_MI = 2019-2010
total_settlement_MI = sum(detroit_data$amount_awarded)
yearly_average_MI = total_settlement_MI/number_years_MI

# Append elements to vector
city = append(city, "Detroit, MI")
no.ofYears = append(no.ofYears, number_years_MI)
Total = append(Total, total_settlement_MI)
YearlyAvg = append(YearlyAvg, yearly_average_MI)
```

## Fort Lauderdale, FL
```{r import_FortLauderdale_Data}
fort_lauderdale_data = read.csv(file = './fort_lauderdale_edited.csv')
```

```{r police_settlement_amts_FortLauderdale}
number_years_FL = 2019-2011 + 1
total_settlement_FL = sum(fort_lauderdale_data$amount_awarded)
yearly_average_FL = total_settlement_FL/number_years_FL

# Append elements to vector
city = append(city, "Fort Lauderdale, FL")
no.ofYears = append(no.ofYears, number_years_FL)
Total = append(Total, total_settlement_FL)
YearlyAvg = append(YearlyAvg, yearly_average_FL)
```

## Indianapolis, IN
```{r import_Indianapolis_Data}
indianapolis_data = read.csv(file = './indianapolis_edited.csv')
```

```{r police_settlement_amts_Indianapolis}
number_years_IN = 2019-2010 + 1
total_settlement_IN = sum(indianapolis_data$amount_awarded)
yearly_average_IN = total_settlement_IN/number_years_IN

# Append elements to vector
city = append(city, "Indianapolis, IN")
no.ofYears = append(no.ofYears, number_years_IN)
Total = append(Total, total_settlement_IN)
YearlyAvg = append(YearlyAvg, yearly_average_IN)
```

## Little Rock, AR
```{r import_LittleRock_Data}
little_rock_data = read.csv(file = './little_rock_edited.csv')
```

```{r police_settlement_amts_LittleRock}
number_years_AR = 2019-2010 + 1
total_settlement_AR = sum(little_rock_data$amount_awarded)
yearly_average_AR = total_settlement_AR/number_years_AR

# Append elements to vector
city = append(city, "Little Rock, AR")
no.ofYears = append(no.ofYears, number_years_AR)
Total = append(Total, total_settlement_AR)
YearlyAvg = append(YearlyAvg, yearly_average_AR)
```

## Los Angeles, CA
```{r import_LosAngeles_Data}
los_angeles_data = read.csv(file = './los_angeles_edited.csv')
```

```{r police_settlement_amts_LosAngeles}
number_years_CA = 2019-2010 + 1
total_settlement_CA = sum(los_angeles_data$amount_awarded)
yearly_average_CA = total_settlement_CA/number_years_CA

# Append elements to vector
city = append(city, "Los Angeles, CA")
no.ofYears = append(no.ofYears, number_years_CA)
Total = append(Total, total_settlement_CA)
YearlyAvg = append(YearlyAvg, yearly_average_CA)
```

## Memphis, TN
```{r import_Memphis_Data}
memphis_data = read.csv(file = './memphis_edited.csv')
```

```{r police_settlement_amts_Memphis}
number_years_TN = 2019-2013 + 1
total_settlement_TN = sum(memphis_data$amount_awarded)
yearly_average_TN = total_settlement_TN/number_years_TN

# Append elements to vector
city = append(city, "Memphis, TN")
no.ofYears = append(no.ofYears, number_years_TN)
Total = append(Total, total_settlement_TN)
YearlyAvg = append(YearlyAvg, yearly_average_TN)
```

## Miami, FL
```{r import_Miami_Data}
miami_data = read.csv(file = './miami_edited.csv')
```

```{r police_settlement_amts_Miami}
number_years_FL = 2020-2010 + 1
total_settlement_FL = sum(miami_data$amount_awarded)
yearly_average_FL = total_settlement_FL/number_years_FL

# Append elements to vector
city = append(city, "Miami, FL")
no.ofYears = append(no.ofYears, number_years_FL)
Total = append(Total, total_settlement_FL)
YearlyAvg = append(YearlyAvg, yearly_average_FL)
```

## Milwaukee, WI
```{r import_Milwaukee_Data}
milwaukee_data = read.csv(file = './milwaukee_edited.csv')
```

```{r police_settlement_amts_Milwaukee}
number_years_WI = 2019-2010 + 1
total_settlement_WI = sum(milwaukee_data$amount_awarded)
yearly_average_WI = total_settlement_WI/number_years_WI

# Append elements to vector
city = append(city, "Milwaukee, WI")
no.ofYears = append(no.ofYears, number_years_WI)
Total = append(Total, total_settlement_WI)
YearlyAvg = append(YearlyAvg, yearly_average_WI)
```

## New Orleans, LA
```{r import_NewOrleans_Data}
new_orleans_data = read.csv(file = './new_orleans_edited.csv')
```

```{r police_settlement_amts_NewOrleans}
number_years_LA = 2019-2010 + 1
total_settlement_LA = sum(new_orleans_data$amount_awarded)
yearly_average_LA = total_settlement_LA/number_years_LA

# Append elements to vector
city = append(city, "New Orleans, LA")
no.ofYears = append(no.ofYears, number_years_LA)
Total = append(Total, total_settlement_LA)
YearlyAvg = append(YearlyAvg, yearly_average_LA)
```

## New York, NY
```{r import_NewYork_Data}
new_york_data = read.csv(file = './new_york_edited.csv')
```

```{r police_settlement_amts_NewYork}
number_years_NY = 2019-2010 + 1
total_settlement_NY = sum(new_york_data$amount_awarded)
yearly_average_NY = total_settlement_NY/number_years_NY

# Append elements to vector
city = append(city, "New York, NY")
no.ofYears = append(no.ofYears, number_years_NY)
Total = append(Total, total_settlement_NY)
YearlyAvg = append(YearlyAvg, yearly_average_NY)
```

## North Charleston, SC
```{r import_NorthCharleston_Data}
north_charleston_data = read.csv(file = './north_charleston_edited.csv')
```

```{r police_settlement_amts_NorthCharleston}
number_years_SC3 = 2019-2010 + 1
total_settlement_SC3 = sum(north_charleston_data$amount_awarded)
yearly_average_SC3 = total_settlement_SC3/number_years_SC3

# Append elements to vector
city = append(city, "North Charleston, SC")
no.ofYears = append(no.ofYears, number_years_SC3)
Total = append(Total, total_settlement_SC3)
YearlyAvg = append(YearlyAvg, yearly_average_SC3)
```

## Orlando, FL
```{r import_Orlando_Data}
orlando_data = read.csv(file = './orlando_edited.csv')
```

```{r police_settlement_amts_Orlando}
number_years_FL2 = 2018-2010 + 1
total_settlement_FL2 = sum(orlando_data$amount_awarded)
yearly_average_FL2 = total_settlement_FL2/number_years_FL2

# Append elements to vector
city = append(city, "Orlando, FL")
no.ofYears = append(no.ofYears, number_years_FL2)
Total = append(Total, total_settlement_FL2)
YearlyAvg = append(YearlyAvg, yearly_average_FL2)
```

## Paterson, NJ
```{r import_Paterson_Data}
paterson_data = read.csv(file = './paterson_edited.csv')
```

```{r police_settlement_amts_Paterson}
number_years_NJ = 2019-2010 + 1
total_settlement_NJ = sum(paterson_data$amount_awarded)
yearly_average_NJ = total_settlement_NJ/number_years_NJ

# Append elements to vector
city = append(city, "Paterson, NJ")
no.ofYears = append(no.ofYears, number_years_NJ)
Total = append(Total, total_settlement_NJ)
YearlyAvg = append(YearlyAvg, yearly_average_NJ)
```

## Philadelphia, PA
```{r import_Philadelphia_Data}
philly_data = read.csv(file = './philly_edited.csv')
```

```{r police_settlement_amts_Philadelphia}
number_years_PA = 2019-2009 + 1
total_settlement_PA = sum(philly_data$amount_awarded)
yearly_average_PA = total_settlement_PA/number_years_PA

# Append elements to vector
city = append(city, "Philadelphia, PA")
no.ofYears = append(no.ofYears, number_years_PA)
Total = append(Total, total_settlement_PA)
YearlyAvg = append(YearlyAvg, yearly_average_PA)
```

## Richmond, VA
```{r import_Richmond_Data}
richmond_data = read.csv(file = './richmond_edited.csv')
```

```{r police_settlement_amts_Richmond}
number_years_VA = 2019-2010 + 1
total_settlement_VA = sum(richmond_data$amount_awarded)
yearly_average_VA = total_settlement_VA/number_years_VA

# Append elements to vector
city = append(city, "Richmond, VA")
no.ofYears = append(no.ofYears, number_years_VA)
Total = append(Total, total_settlement_VA)
YearlyAvg = append(YearlyAvg, yearly_average_VA)
```

## Roanoke, VA
```{r import_Roanoke_Data}
roanoke_data = read.csv(file = './roanoke_edited.csv')
```

```{r police_settlement_amts_Roanoke}
number_years_VA2 = 2019-2010 + 1
total_settlement_VA2 = sum(roanoke_data$amount_awarded)
yearly_average_VA2 = total_settlement_VA2/number_years_VA2

# Append elements to vector
city = append(city, "Roanoke, VA")
no.ofYears = append(no.ofYears, number_years_VA2)
Total = append(Total, total_settlement_VA2)
YearlyAvg = append(YearlyAvg, yearly_average_VA2)
```

## San Francisco, CA
```{r import_SanFrancisco_Data}
san_francisco_data = read.csv(file = './san_francisco_edited.csv')
```

```{r police_settlement_amts_SanFrancisco}
number_years_CA2 = 2019-2010 + 1
total_settlement_CA2 = sum(san_francisco_data$amount_awarded)
yearly_average_CA2 = total_settlement_CA2/number_years_CA2

# Append elements to vector
city = append(city, "San Francisco, CA")
no.ofYears = append(no.ofYears, number_years_CA2)
Total = append(Total, total_settlement_CA2)
YearlyAvg = append(YearlyAvg, yearly_average_CA2)
```

## Springfield, MA
```{r import_Springfield_Data}
springfield_data = read.csv(file = './springfield_edited.csv')
```

```{r police_settlement_amts_Springfield}
number_years_MA3 = 2020-2006 + 1
total_settlement_MA3 = sum(springfield_data$amount_awarded)
yearly_average_MA3 = total_settlement_MA3/number_years_MA3

# Append elements to vector
city = append(city, "Springfield, MA")
no.ofYears = append(no.ofYears, number_years_MA3)
Total = append(Total, total_settlement_MA3)
YearlyAvg = append(YearlyAvg, yearly_average_MA3)
```

## St Louis, MO
```{r import_StLouis_Data}
stlouis_data = read.csv(file = './stlouis_edited.csv')
```

```{r police_settlement_amts_StLouis}
number_years_MO = 2019-2015 + 1
total_settlement_MO = sum(stlouis_data$amount_awarded)
yearly_average_MO = total_settlement_MO/number_years_MO

# Append elements to vector
city = append(city, "St Louis, MO")
no.ofYears = append(no.ofYears, number_years_MO)
Total = append(Total, total_settlement_MO)
YearlyAvg = append(YearlyAvg, yearly_average_MO)
```

## Washington, DC
```{r import_Washington_Data}
DC_data = read.csv(file = './DC_edited.csv')
```

```{r police_settlement_amts_Washington}
number_years_DC = 2019-2010 + 1
total_settlement_DC = sum(DC_data$amount_awarded)
yearly_average_DC = total_settlement_DC/number_years_DC

# Append elements to vector
city = append(city, "Washington, DC")
no.ofYears = append(no.ofYears, number_years_DC)
Total = append(Total, total_settlement_DC)
YearlyAvg = append(YearlyAvg, yearly_average_DC)
```

## Waterbury, CT
```{r import_Waterbury_Data}
waterbury_data = read.csv(file = './waterbury_edited.csv')
```

```{r police_settlement_amts_Waterbury}
number_years_CT = 2019-2011 + 1
total_settlement_CT = sum(waterbury_data$amount_awarded)
yearly_average_CT = total_settlement_CT/number_years_CT

# Append elements to vector
city = append(city, "Waterbury, CT")
no.ofYears = append(no.ofYears, number_years_CT)
Total = append(Total, total_settlement_CT)
YearlyAvg = append(YearlyAvg, yearly_average_CT)
```

## Figure 1
```{r figure1}
df <- data.frame(city,no.ofYears,Total,YearlyAvg)
df
```

## Figure 2
```{r figure2}
# Dataframe for settlement amounts per year
df1 <- cleveland_data %>% 
  group_by(calendar_year) %>%
  summarise(sum_settlement = sum(amount_awarded)) %>%
  filter(calendar_year != 2020)
df1

# Change calendar_year variable from discrete variable to continuous variable by using factor function
ggplot(df1, aes(x=factor(calendar_year), y = sum_settlement)) + geom_bar(stat = "identity", fill="darkmagenta") +
  xlab("Years") +
  ylab("Total amount paid in settlements") +
  ggtitle("Cleveland's settlement amounts rose after Rice's death")
```

